Pool Model Paper

### 2023-10-26 pool_model.tex
species inhibition: add plots

\subsubsection{Species Inhibition}
\cmt{\cite{lenski_coexistence_1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}
The simplest way to introduce inhibition of species A by species B is via a growth rate of species A depending on the abundance of species B. 
For sake of simplicity we ignore the lag pool and write for a mutual inhibition:
\begin{eqnarray}
\dot{G}_A &=& \mathcal{B}_A(G_B)G_A\left(1 - \frac{G_A+G_B}{N_t}\right )\\
\dot{G}_B &=& \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right ). 
\end{eqnarray}
The function $\mathcal{B}_{A/B}$ captures the inhibiting effects in an effective way. 
It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a waste pool. 
One possible choice for $\mathcal{B}$ is: 
\begin{eqnarray}
\mathcal{B}(G)&=&\frac{\alpha}{1+KG}. 
\end{eqnarray}
This corresponds to the case $W\sim G$, where $W$ is the waste (or whatever one chooses to call it) of species A affecting species B. 
If the waste accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:
\begin{eqnarray}
\dot{G}_A &=& \frac{\alpha_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right )\\
\dot{G}_B &=& \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right ). 
\end{eqnarray}
One can distinguish two limiting cases: {\it i}) $K_{A/B}\ll N_t$ and {\it ii}) $K_{A/B}\gg N_t$. 
In the first case the inhibition is of minor importance and could be ignored. 
In the latter case one can simplify the equations, given $n_A$ and $n_B$ are non-zero:
\begin{eqnarray}
\label{Mut_Inhib}
\dot{G}_A &=&\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right )\\
\dot{G}_B &=& \tilde{\alpha}_B\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
\end{eqnarray}
We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$. Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$. Rescaling time with $\tilde{\alpha}_A$ and defining the dimensionless parameter 
\begin{eqnarray}
\psi&=&\tilde{\alpha}_B/\tilde{\alpha}_A\\
\psi&=&\frac{K_B\alpha_B}{K_A\alpha_A}
\end{eqnarray}
results in: 
\begin{eqnarray}
\dot{G}_A &=&\frac{G_A}{G_B}\left(1 - G_A-G_B\right )\\
\dot{G}_B &=&\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
\end{eqnarray}
We can separate three cases: 
\begin{eqnarray}
\psi&>&1:\quad\text{Species B growth faster and/or inhibits A stronger}\\
\psi&=&1:\quad\text{A and B are equivalent}\\
\psi&<&1:\quad\text{Species A growth faster and/or inhibits B stronger}
\end{eqnarray}

>>>>

\subsubsection{Species Inhibition}
\cmt{\cite{lenski_coexistence_1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}
The simplest way to introduce inhibition of species A by species B is via a growth rate of species A depending on the abundance of species B. 
For sake of simplicity we ignore the lag pool and write for a mutual inhibition:
\begin{eqnarray}
\dot{G}_A &=& \mathcal{B}_A(G_B)G_A\left(1 - \frac{G_A+G_B}{N_t}\right )\\
\dot{G}_B &=& \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right ). 
\end{eqnarray}
The function $\mathcal{B}_{A/B}$ captures the inhibiting effects in an effective way. 
It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a waste pool. 
One possible choice for $\mathcal{B}$ is: 
\begin{eqnarray}
\mathcal{B}(G)&=&\frac{\alpha}{1+KG}. 
\end{eqnarray}
This corresponds to the case $W\sim G$, where $W$ is the waste (or whatever one chooses to call it) of species A affecting species B. 
If the waste accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:
\begin{eqnarray}
\dot{G}_A &=& \frac{\alpha_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right )\\
\dot{G}_B &=& \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right ). 
\end{eqnarray}
One can distinguish two limiting cases: {\it i}) $K_{A/B}\ll N_t$ and {\it ii}) $K_{A/B}\gg N_t$. 
In the first case the inhibition is of minor importance and could be ignored. 
In the latter case one can simplify the equations, given $n_A$ and $n_B$ are non-zero:
\begin{eqnarray}
\label{Mut_Inhib}
\dot{G}_A &=&\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right )\\
\dot{G}_B &=& \tilde{\alpha}_B\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
\end{eqnarray}
We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$. Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$. Rescaling time with $\tilde{\alpha}_A$ and defining the dimensionless parameter 
\begin{eqnarray}
\psi&=&\tilde{\alpha}_B/\tilde{\alpha}_A\\
\psi&=&\frac{K_B\alpha_B}{K_A\alpha_A}
\end{eqnarray}
results in: 
\begin{eqnarray}
\dot{G}_A &=&\frac{G_A}{G_B}\left(1 - G_A-G_B\right )
\label{eq:2sp_1pool_inhibit_unitless1}\\
\dot{G}_B &=&\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
\label{eq:2sp_1pool_inhibit_unitless2}
\end{eqnarray}
We can separate three cases: 
\begin{eqnarray}
\psi&>&1:\quad\text{Species B growth faster and/or inhibits A stronger}\\
\psi&=&1:\quad\text{A and B are equivalent}\\
\psi&<&1:\quad\text{Species A growth faster and/or inhibits B stronger}
\end{eqnarray}

\begin{figure}[H]
    \begin{center}
    \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}
    \caption{{\footnotesize The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
    Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
    The relation between inhibition strength of spesies $B$ and $A$ are defined by controll parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.}}
    \label{fig:1pool_2sp_inhibit}
    \end{center}
\end{figure}
